Towards a Cognitive System That Can Recognize Spatial Regions Based on Context

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex

Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability. If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand and refer to the same spatial regions humans do. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc. In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm.

\from{klenk}{Cite recent Langley work}

@@

We use \textit{anchor points}~\cite{Klenk/etal2005} to define the boundaries of CDSRs. Anchor points are symbolic descriptions which link a conceptual entity to a perceived entity. The perceived entities we use are the objects recognised by Dora, and the room itself. The room representation is created by putting a convex hull around the lines in Dora's SLAM map. Anchor points are created from perceived entities using unary functions, e.g. \fw{(XMaxYMostFn Desk1)} represents the point on the \fw{Desk1} with the largest x coordinate taken from the set of points with a y coordinate within 5\% of the maximum y coordinate. Anchor points are linked to particular CDSRs using a \fw{regionBoundary} relation. After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation. Therefore, each CDSR has one type and a variable number of boundary segments.

@@

\fw{
(regionType CDSR9 FrontRegion) \\
(regionBoundarySegment CDSR9 \\
\hspace*{2em}(YMaxXFewestFn Room-Base) \\
\hspace*{2em}(YMinXFewestFn Room-Base)) \\
(regionBoundarySegment CDSR9 \\
\hspace*{2em}(YMinXFewestFn Room-Base) \\
\hspace*{2em}(YMinXFewestFn Group1))\\
...
}
}
  \caption{Representing the front of the classroom context-dependent region \fw{CDSR9}}
  \label{fig:cdsr-reps}
\end{figure}

Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room1} which is pictured in the top of Figure~\ref{fig:rooms}. The anchor points (connected in orange in Figure~\ref{fig:rooms}) define the region. \fw{(LeftmostBottomFn Room1)} and \fw{(LeftmostTopFn Room1)} are the points with the lowest x coordinate out of the set of points defined as the top and bottom of \fw{Room1} respectively. The sets of points representing the top and bottom of the classroom are found by selecting the points defining the area within a threshold of the area's maximum or minimum y coordinate. The other corners of the context-dependent spatial region are defined by the four desks and the whiteboard. Because perceived objects (e.g. the desks) are defined as single points, they can serve as anchor points directly. The semantic label \fw{FrontRegion} ties the geometric extent of the boundary points to a conceptual region. This definition for the front of the classroom is specific to \fw{Room1} and its entities. It is clearly context-dependent because its extent is dependent on the arrangement of the anchor points used to define its boundary. If the desks were in a different position then the region would cover a different extent (e.g. if they were closer to the whiteboard then the region would be smaller).

@@

% \from{Nick}{This section feels a bit long. Can we cut it down at all? Perhaps the second half of the second paragraph if we don't use the constraints explicitly}
% done

We assume that a cognitive system will have a way of initially acquiring examples of CDSRs, e.g., by being taught through dialogue, sketching, or hand-coding. To avoid burdening potential tutors with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training. For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building. We have chosen \emph{analogy} as the approach our system will use to solve this problem, as analogy has been previously used to successfully combine semantic as well as geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.

Analogy is an essential cognitive process. In humans, analogical processing has been observed in language comprehension, problem-solving, and generalization \cite{Gentner2003}. The structure-mapping theory of analogy and similarity postulates this process as an alignment between two structured representations, a \textit{base} and a \textit{target} \cite{Gentner1983a}. 
% This alignment process is governed by three constraints: \textit{identicality}, \textit{parallel connectivity}, and \textit{one-to-one mapping}. The identicality constraint provides a strong preference for only allowing identical predicates to match. Parallel connectivity states that if two predicates are matched then their arguments must also match. The one-to-one mapping constraint requires that each element in the base corresponds to at most one element in the target and vice versa. To select between competing mappings, the \textit{systematicity} principle prefers mappings that are highly interconnected and contain deep chains of higher order relations over mappings with an equal number of relations which are independent from each other.
We use the Structure-Mapping Engine (SME) \cite{Falkenhainer1989a} to perform analogical matching in our system. Given base and target representations as input, SME produces one or more mappings. Each mapping is represented by a set of \textit{correspondences} between entities and expressions in the base and target structures.  Mappings also include \textit{candidate inferences} which are conjectures about the target using expressions from the base which, while unmapped in their entirety, have subcomponents that participate in the mapping's correspondences. SME operates in polynomial time, using a greedy algorithm \cite{Forbus/etal1994}.

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{images/AnalogyDiagram.png}
  \caption{Analogical mapping between five base expressions and three target expressions. }
  %Ovals and squares represent relations and entities respectively. Black dashed bi-directional arrows are the correspondences between the base and target, and the red dashed objects represent a candidate inference projected into the target.

@@

In our system, the base and target representations consist of the entities Dora has perceived in two different rooms, plus the QSRs computed on top of these. The base also contains a labeled CDSR which we assume is sought in target. The result of result of running SME on these representations is a set of correspondences between the base and target, and a set of candidate inferences about the target. We use these to transfer the CDSR from base to target (i.e. recognizing the CDSR in the target) as follows. First, we identify the CDSR of the sought type in the base. This entity should not participate in the mapping, but should be included in \fw{:skolem} expressions in the candidate inferences. To transfer the CDSR to the target, we collect the candidate inferences that result from \fw{regionBoundary} statements mentioning the base CDSR. The second arguments of these candidate inferences are anchor points in the target environment. We use these to define the boundary of the CDSR in the target, linking them to a new CDSR entity using \fw{regionBoundary} statements.

@@

We first create the base and target representations by running Dora in the two different classrooms. These are pictured in the top and bottom of Figure~\ref{fig:rooms} respectively. In each case, Dora is manually driven around the room to allow it to create a metric map. Once the map is created, Dora is then positioned such that the objects are observable and the visual recognition system is run. The map and object data that results from this is then passed on to the QSR generator. In the base case Dora perceives 2 desks, a whiteboard and the room area. To this we add the CDSR representing the front of the room. The case includes a total of \textbf{SOMENUMBER} expressions relating the \textbf{SOMENUMBER} entities. \textbf{SOMENUMBER} of these expressions are used to define the boundary of a CDSR representing the front of the room: the \textbf{SOMENUMBER} desks in the front row, the whiteboard, and the leftmost bottom and leftmost top of the room. The target case includes \textbf{SOMENUMBER}. The target case includes \textbf{SOMENUMBER} statements and \textbf{SOMENUMBER} entities.

SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation. The resulting analogy includes \textbf{SOMENUMBER} correspondences between the entities and expressions in the base and target and \textbf{SOMENUMBER} candidate inferences.  All \textbf{SOMENUMBER} \fw{regionBoundary} statements in the base appear as candidate inferences. \textbf{SOMENUMBER} of the anchor points are exactly correct, the \textbf{\dots}. The other \textbf{SOMENUMBER} analogically transferred anchor points are to the \textbf{\dots}. The green region in the lower image of Figure \ref{fig:rooms} illustrates the CDSR created by the \textbf{SOMENUMBER} boundary points in the target environment.

@@

In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions. The system is able to use analogy to recognise CDSRs in previously unseen environments through a process of analogy with a manually annotated known room. The system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy. In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with in future work. These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g. MAC/FAC \cite{forbus/etal1995}), and a lack of transfer post-processing (e.g. comparing the QSRs present in both base and transferred regions) to improve results. Despite these limitations our evaluation demonstrated that the system is able to match the CDSRs produced by trial subjects on unseen rooms to an acceptable degree \textbf{SOMENUMBER}.

>>>>

Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability. If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand and refer to the same spatial regions humans do. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc. In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm \cite{langley:inpress}.

@@

We use \textit{anchor points}~\cite{Klenk/etal2005} to define the boundaries of CDSRs. Anchor points are symbolic descriptions which link a conceptual entity to a perceived entity. The perceived entities we use are the objects recognised by Dora, and the room itself. The room representation is created by putting a convex hull around the lines in Dora's SLAM map. Anchor points are created from perceived entities using unary functions, e.g. \fw{(XMaxYMostFn Desk1)} represents the point on the \fw{Desk1} with the largest x coordinate taken from the set of points with a y coordinate within 5\% of the maximum y coordinate. Anchor points are linked to particular CDSRs using a \fw{boundarySegment} ternary relation. After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation. Therefore, each CDSR has one type and a variable number of boundary segments.

@@

\fw{
(regionType CDSR9 FrontRegion) \\
(boundarySegment CDSR9 \\
\hspace*{2em}(YMaxXFewestFn Room3) \\
\hspace*{2em}(YMinXFewestFn Room3)) \\
(boundarySegment CDSR9 \\
\hspace*{2em}(YMinXFewestFn Room3) \\
\hspace*{2em}(YMinXFewestFn Group1))\\
...
}
}
  \caption{Representing the front of the classroom context-dependent region \fw{CDSR9}}
  \label{fig:cdsr-reps}
\end{figure}

Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room3} which is pictured in the top of Figure~\ref{fig:rooms}. The boundary segments (shown in orange in Figure~\ref{fig:rooms}) define the extent of the region. \fw{(YMaxXFewestFn Room3)} and \fw{(YMinXFewestFn Room3)} are the points with the highest and lowest y coordinate out of the set of points within 5\% of the minimum x coordinate of \fw{Room3}. The next segment connect the lower left coordinate in the figure to the \fw{(YMinXFewestFn Group1)}, where \fw{Group1} includes the eight desks. There are two more boundary segments completing a polygon for this region. The semantic label \fw{FrontRegion} ties this polygon to a conceptual region, "the front of the room". This definition for the front of the room is specific to \fw{Room3} and its entities. It is clearly context-dependent because its extent is dependent on the arrangement of the anchor points used to define its boundary. If the desks were in a different position then the region would cover a different extent (e.g. if they were further to the left then the region would be smaller).

@@

We assume that a cognitive system will have a way of initially acquiring examples of CDSRs, e.g., by being taught through dialogue, sketching, or hand-coding. To avoid burdening potential users with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training. For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building. We have chosen \emph{analogy} as the approach our system will use to solve this problem, as analogy has been previously used to successfully combine semantic as well as geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.

Analogy is an essential cognitive process. In humans, analogical processing has been observed in language comprehension, problem-solving, and generalization \cite{Gentner2003}. The structure-mapping theory of analogy and similarity postulates this process as an alignment between two structured representations, a \textit{base} and a \textit{target} \cite{Gentner1983a}. 
% This alignment process is governed by three constraints: \textit{identicality}, \textit{parallel connectivity}, and \textit{one-to-one mapping}. The identicality constraint provides a strong preference for only allowing identical predicates to match. Parallel connectivity states that if two predicates are matched then their arguments must also match. The one-to-one mapping constraint requires that each element in the base corresponds to at most one element in the target and vice versa. To select between competing mappings, the \textit{systematicity} principle prefers mappings that are highly interconnected and contain deep chains of higher order relations over mappings with an equal number of relations which are independent from each other.
We use the Structure-Mapping Engine (SME) \cite{Falkenhainer1989a} to perform analogical matching in our system. Given base and target representations as input, SME produces one or more mappings. Each mapping is represented by a set of \textit{correspondences} between entities and expressions in the base and target structures. Mappings are defined by expressions with an identical relation and corresponding arguments. When provided with expression strengths, such as, our spatial relationships, SME prefers mappings with closely aligned fact strengths. SME can be given \textit{pragmatic constraints} that require certain entities in the base to be included in the mapping. Mappings also include \textit{candidate inferences} which are conjectures about the target using expressions from the base which, while unmapped in their entirety, have subcomponents that participate in the mapping's correspondences. SME operates in polynomial time, using a greedy algorithm \cite{Forbus/etal1994}.

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{images/AnalogyDiagram.png}
  \caption{Analogical mapping between six base expressions and three target expressions. }
  %Ovals and squares represent relations and entities respectively. Black dashed bi-directional arrows are the correspondences between the base and target, and the red dashed objects represent a candidate inference projected into the target.

@@

In our system, the base and target representations consist of the entities Dora has perceived in two different rooms, the QSRs between them and any groups that have been identified. The base also contains a labeled CDSR of the type sought in target. The result of running SME on these representations is a set of correspondences between the base and target, and a set of candidate inferences about the target. We use these to transfer the CDSR from base to target (i.e. recognizing the CDSR in the target) as follows. First, we identify the CDSR of the sought type in the base and use SME's pragmatic constraints to ensure that the entities referred to its anchor points participate in the mapping. To transfer the CDSR to the target, we collect the candidate inferences that result from \fw{boundarySegment} statements mentioning the base CDSR. The second and third arguments of these candidate inferences are anchor points in the target environment. We use these to define the boundary of the CDSR in the target.

@@

We first create the base and target representations by running Dora in the two different classrooms. These are pictured in the top and bottom of Figure~\ref{fig:rooms} respectively. In each case, Dora is manually driven around the room to allow it to create a metric map. Once the map is created, Dora is then positioned such that the objects are observable and the visual recognition system is run. The map and object data that results from this is then passed on to the QSR generator. In the base case, Dora perceives 8 individual desks, a group entity containing these desks and the room area. To this we add the CDSR representing the front of the room. The case includes a total of 50 expression relating the 20 entities. Six of these expressions are used to define the boundary segments and CDSR representing the front of the room. The target case includes 26 expressions and 11 entities.

SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation requiring \fw{Room3} and \fw{Group1} participate in the mapping as they are referenced by the anchor points in the base. The resulting analogy includes 26 correspondences between the entities and expressions and 32 candidate inferences. Four of these candidate inferences define the CDSR in the target with anchor points to the group of desks and the room in the target. The green region in the lower image of Figure \ref{fig:rooms} illustrates the inferred CDSR.

@@

In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions. The system identifies CDSRs in previously unseen environments through analogy with a manually annotated known room. The system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy. In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with future work. These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g. MAC/FAC \cite{forbus/etal1995}), and a lack of transfer post-processing (e.g. comparing the QSRs present in both base and transferred regions) to improve results. Despite these limitations our evaluation demonstrated that the system is able to match the CDSRs produced by trial subjects on unseen rooms to an acceptable degree \textbf{SOMENUMBER}.
